### 2021-04-30 Paper_replication_files.Rmd
creating data folder

---
  title: "Replication file for the paper: A multidimensional feasibility evaluation of low-carbon scenarios"
output:
  html_document: default
always_allow_html: true
---

  ```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.par = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r packages, include=FALSE, MESSAGE=FALSE}
pkgs<-c("vroom", "tidyverse", "plotly", "RColorBrewer", "viridis", "hrbrthemes", "here", "ggridges", "ggsci", "ggrepel", "stringr", "cowplot", "ggfan", "scales", "skimr", "jtools", "arsenal", "corrplot", "countrycode", "ggpmisc")
#install.packages(pkgs) -> please uncomment it if you don't have the required packages installed

load<-lapply(pkgs, library, character.only=TRUE)
```

```{r functions, include=FALSE, MESSAGE=FALSE}

geom_mean<-function(x){exp(mean(log(x)))}

```

```{r data_input, include=FALSE, MESSAGE=FALSE}

#Data from the 1.5C scenario explorer
world<-vroom(here("input", "full.csv"))
meta<-vroom(here("input", "meta.csv"))
marker_set<-vroom(here("input", "marker.csv"))
meta_policy<-vroom(here("input", "meta_policy_1.csv"))
baselines<-vroom(here("input", "baselines.csv"))

#assuptions about thresholds

thresholds<-vroom(here("input", "thresholds1.csv"))

#Regional conversions for R5 for the empirical data that is used
region_con<-vroom(here("input", "region_conversion.csv"))

#Data on coal from Global Monitor

coal_50_all<-vroom(here("input", "coal_50_all.csv"))

#Data from Adrijevic et al.(2019, 2020) and the Wittgenstein Center

governance_proj<-vroom(here("input", "governance.csv"))
gov_eff<-vroom(here("input", "gov_eff_projections.csv"))
GII<-vroom(here("input", "GII.csv"))
EPI<-vroom(here("input", "EPI.csv"))
gov_his<-vroom(here("input", "gov_historical.csv"))
age<-vroom(here("input", "age1.csv"))
education<-vroom(here("input", "education1.csv"))
fertility<-vroom(here("input", "fertility1.csv"))
```